---
title: "R codes for Ch 18"
author: "Dr. Hasthika Rupasinghe"
date: '`r format(Sys.time(), "%b %d, %Y at %X")`'
output:
  bookdown::html_document2:
    highlight: textmate
    theme: yeti
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = "center", comment = NA)
#library(tidyverse)
```

# Testing Hypotheses about proportions


#### Step 1:  Formulate Hypotheses (PARAMETERS) {-}

```{r}
#Step 1:  Formulate Hypotheses (PARAMETERS)
# Assign the value from H0 to p
p0 <- .9
```


#### Step 2:  Calcualte the standardized test statistic {-}
```{r}
n <- 400
phat <- 330/400 #or a percent if given

zstat <- (phat - p0)/sqrt(p0*(1-p0)/n)

cat("The standardized test statistic is",zstat)
```

#### Step 3:  Calcualte the P-value {-}

```{r}
#P-value = P(get what you got or more when Ho is true)

#less than
LTPval <- pnorm(zstat)
#greater than
GTPval <- 1-pnorm(zstat)
#not equal
NEPval <- 2*pnorm(-abs(zstat))

cat("P-value is")
list(Less=LTPval,Greater=GTPval,NotEqual=NEPval)

```

#### Step 4: Conclusion {-}
```{r}
#Statistical conclusion: A) Reject Ho if P-val < alpha (use .05 if none given)
#                         B) Fail to reject Ho if P-val > alpha

#English conclusion: A) There IS evidence of Ha (written to match problem)
#                     B) There is NOT evidence of Ha.

# End of Hypotheses about proportions



```

## Checking assumptions 

```{r}
# Check assumptions

#Independence
#Randomization
#Less than 10% of the population in the sample
#Success/Failure

n <- 400

n*p0 > 10 
n*(1-p0) >10
#need this to be TRUE TRUE to continue
```


# Testing Hypotheses about mean $\mu$

#### Step 0:  Find(or enter) the sample mean ($\bar{y}$) and standard deviation ($s$) and $n${-}

```{r}
ybar <- 12.5
s <- 2

n <- 25
```


#### Step 1:  Formulate Hypotheses (PARAMETERS) {-}

```{r}
#Step 1:  Formulate Hypotheses (PARAMETERS)

mu0 <- 10 #This is the value in H0 : mu = mu0
```


#### Step 2:  Calcualte the standardized test statistic {-}
```{r}
# Step 2:  Calcualte the standardized test statistic
tstat <- 1.76 #(ybar - mu0)/(s/sqrt(n))

cat("The t-stat is",tstat)
```

#### Step 3:  Calcualte the P-value {-}

```{r}
#P-value = P(get what you got or more when Ho is true)

df <- 24 #n - 1 # degrees of freedom

#less
LTPval<- pt(tstat,df)

#more
GTPval<- 1-pt(tstat,df)

#not equal
NEPval <- 2*pt(-abs(tstat), df)

list(Less=LTPval,Greater=GTPval,NotEqual=NEPval)

```

#### Step 4: Conclusion {-}
```{r}
#Statistical conclusion: A) Reject Ho if P-val < alpha (use .05 if none given)
#                         B) Fail to reject Ho if P-val > alpha

#English conclusion: A) There IS evidence of Ha (written to match problem)
#                     B) There is NOT evidence of Ha.

# End of Hypotheses about the mean
```


## Sample size to estimate $\mu$ within a given ME

You write it!

```{r eval=FALSE, include=FALSE}
CIlev <- .95
s <- 29.31
ME <- 0.02

alpha <- 1-CIlev
alphaovertwo <- alpha/2
oneminusalphaovertwo <- 1- alpha/2
zinCI <- qnorm(oneminusalphaovertwo)

nprelim <- (zinCI * s/ME)^2
df <- nprelim-1

tstar <- qt(oneminusalphaovertwo,df)

nfinal <- (tstar * s/ME)^2
ceiling(nfinal)
```





