---
title: "R codes for Ch 20 PART II: Two Means"
author: "Dr. Hasthika Rupasinghe"
date: '`r format(Sys.time(), "%b %d, %Y at %X")`'
output:
  bookdown::html_document2:
    toc: true
    highlight: textmate
    theme: yeti
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = "center", comment = NA)
#library(tidyverse)
```

> **_NOTE:_**  Use this script for Ch20, ONLY if the question talks about **two means**

# Checking assumptions

  1) Independent data
  2) Randomization
  3) Less than 10\% of the population is in the sample
  5) Independent Groups

If assumptions are met, $\bar{x_1} - \bar{x_2} \sim N(\mu_1-\mu_2, \sqrt{(s_1^2/n_1 +s_2^2/n_2)})$

# CI for $\mu_1-\mu_2$

You wrote it!

```{r}
ybar1 <- 28.1
s1 <- 3.93
n1 <- 31

ybar2 <- 25.98
s2 <- 2.79
n2 <- 32

df <- (s1^2/n1 +s2^2/n2)^2/(1/(n1-1)*(s1^2/n1)^2 + 1/(n2-1)*(s2^2/n2)^2)

CIlev <- .95

alpha <- 1-CIlev
alphaovertwo <- alpha/2
oneminusalphaovertwo <- 1- alpha/2
tstar <- qt(oneminusalphaovertwo,df)

tstar  #tstar in a CI formula

LowP <- (ybar1-ybar2) - tstar * sqrt(s1^2/n1 +s2^2/n2)
UpP <- ybar1-ybar2 + tstar * sqrt(s1^2/n1 +s2^2/n2)

CI <- c(LowP,UpP)

cat("The ", CIlev*100, " % confidence interval for mu is", CI)
  
```

# HT for $\mu_1-\mu_2$

```{r}
ybar1 <- 45
s1 <- 20
n1 <- 81

ybar2 <- 52
s2 <- 18
n2 <- 77

df <- (s1^2/n1 +s2^2/n2)^2/(1/(n1-1)*(s1^2/n1)^2 + 1/(n2-1)*(s2^2/n2)^2)

cat("Degees of freedom is", df, "\n")

tstat <- (ybar1-ybar2)/sqrt(s1^2/n1 +s2^2/n2)

cat("The t-stat is", tstat, "\n")

#less
LTPval<- pt(tstat, df)

#more
GTPval<- 1-pt(tstat, df)

#not equal
NEPval<-2*pt(-abs(tstat), df)


list(LessPval=LTPval,GreaterPval=GTPval,NotEqualPval=NEPval)
```

# If you had data insted of the summary statitics:

Change the top part of the script (CI or HT) as shown bellow to get the summary statistics.  

```{r}
heat2 <- c(105.9,106.4,106.4,106.5,107,108.8,109.6,115.3)
heat7 <- c(106.1,106.1,106.4,106.9,108.4,108.5,117.5)

ybar1 <- mean(heat2)
s1 <- sd(heat2) 
n1 <- length(heat2)

ybar2 <- mean(heat7)
s2 <- sd(heat7) 
n2 <- length(heat7)
```

