---
title: "R codes for Ch 26: ANOVA"
author: "Dr. Hasthika Rupasinghe"
date: '`r format(Sys.time(), "%b %d, %Y at %X")`'
output:
  bookdown::html_document2:
    toc: true
    highlight: textmate
    theme: yeti
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = "center", comment = NA)
library(tidyverse)
```


 
# ANOVA example

```{r}
ACT <- read.table("https://hasthika.github.io/STT3820/Lecture%20Notes/ScoreACT.txt", sep ="" , header = TRUE)

ACT

act <- stack(ACT)
act

anovaACT<- aov(act$values~act$ind)
outputanovaACT <- anova(anovaACT)
outputanovaACT


```

 <a href="#top">Back to top</a>


# Assumptions check:

## Independence - Check for randomization {-}

## Equal variance among the treatments - Check using Side by side Boxplots {-}

### Side by side Boxplots {-}

```{r}
boxplot(values~ind,  # Specify the variables to graph, quantitative variable first
        data = act,                    # Specify the data set 
        main = "ACT Scores and Groups",           # Create the chart title
        xlab = "Groups (or Treatments)",  # Create the x-axis label
        ylab = "ACT Scores")           # Create the y-axis label

```

## Normal Errors {-}

### Residual plot (Check for outliers) {-}
### Use a normal probability plot (qqplot) of the residuals or a histogram of the resduals {-}
    
```{r}
plot(anovaACT) # Use only the first two plots (Residual plot and qqplot)
```
    




`


# Multiple comparisons (**After** you reject $H_0$)


```{r}
TukeyHSD(aov(act$values~act$ind), conf.level=.95)
```


# HW help

```{r}

# HW#2, #7

library(readxl)
data7 <- read_excel("~/git_repositories/STT3820/Lecture Notes/data7.xlsx")

data7 #see original

data7 <- stack(data7)
data7 #see numerical column and names column

anovadata <- aov(data7$values~data7$ind)
anovadata
summary(anovadata)
```


# P-value from a $F$ Distribution

```{r}
Fstat <- 13.55

numDf <- 3 # Treatment df
denDf <- 8 # Error (or residual) df

1-pf(Fstat, numDf, denDf)

```



